Exploring the Motives and Driving Forces Behind Modifications to APIs From a Software Evolution Perspective

### 2015-05-06 thesis/paper.tex
removed comments
Reviewed disc rq1, added comments

@@

After cross-examining our results with the interviewee and their responses, we found that roughly 78\% of our categorisation matched the actual categorisation of the changes.

%%TODO Finish this off XXX
%New additions to the API were the most frequent change that we observed. Each addition added new functionality to the API, ranging from controlling the camera, enabling audio analysis, allowing external storage connections and providing a serial port to interface with other devices.

%XXX
%These changes were deemed to be modifications or re-implementations of existing functionality of the API. During the interview it was mentioned that one intent behind the re-implementations was to bring them in line with the newly created functionality which takes advantage of Glib, a new kind of error handling and extensive use of callbacks.

%XXX
%While the main focus of our investigation of the source code and documentation was to identify changes between the versions, it was also to discover differences larger trends of the API. The trends were also analysed and given an appropriate maintenance category.

%XXX
%While the changes between versions 1.4 and 2.0 included many deprecations, only one of them was not re-implemented in a new part of the API.

\subsubsection{axptz}
%Add info about this is one of the added changes XXX (do that for all)
The Pan, Tilt and Zoom (ptz) module concerns functionality related to manoeuvring the camera, both mechanically and digitally.

%%Relate it to Categories
We concluded that this change was enhancive as it is a brand-new feature of the API. There was no secondary motive behind this change, as the change was purely enhancive.

%%What she said
The interviewee described how this change and the other changes classified as additions were all included in the new API at the request of API users. Prior to the change being introduced, there was no API functionality for controlling the camera and as such, it was a high priority addition to the new API.

%%Category based on her information
Based on the interviewees' comments, we concluded that we had correctly identified the change as being enhancive. Furthermore, we were also correct in our prediction that there was no secondary category.

\subsubsection{axaudio}

%%Describe the change 
The most important functionality of the camera is to capture and analyse video data, however, some the API users require the possibility to perform analysis upon audio data. The addition of axaudio allows API users to perform audio analysis upon audio streams captured by cameras.

%%Relate it to Categories
This change was determined to be enhancive, as it was an addition of a new feature to the API. We found no secondary motive of the change.

%%What she said
Audio analysis had not previously been possible on Axis devices. During the interview, it was revealed that an API user had managed to implement their own audio analysis using version 1.4 of the API. Given the demand for audio analysis, it was included in version 2.0 of the API.

%%Category based on her information
Based on the the interviewee's reasoning, we concluded that we had successfully determined that the change was enhancive. The addition of axaudio had no secondary motive.

@@

%%Describe the change 
Prior to the newest version of the API being released, the sole alternative for storage was to use the Video Management System. Version 2.0 included the possibility of alternative storage solutions, by allowing to the camera to interface with external storage devices, such as SD cards.

%Relate it to categories
Enhancive was chosen to be the most applicable category of change, given that axstorage was a new addition in version 2.0. No secondary motive was identified.

%%What she said
The limited storage options for the cameras was the reason that increased functionality was strongly requested by API users and as such axstorage was included in version 2.0 of the API.

%%Category based on her information
The information gathered from the interview supported that we were correct in our categorisation of this change. No secondary category was established.

@@

%%Describe the change 
The addition of axserialport allows API users to connect various devices to their cameras by through the serial port. This allows users to integrate multiple devices to create their own unique solutions. In version 1.4 of the API, there was no possibility to connect external devices to the camera.

%%Relate it to Categories
The addition of this module was determined to be an enhancive change, as it involved the addition of a new feature. No relevant secondary category were identified.

%%What she said
The addition of axserialport was requested by API users. A possible use case, which was given during the interview involved a ticketing system which allows for joint analysis of data from both devices. As such it was included in version 2.0, to meet the demand of the API users.

%%Category based on her information
Given this information, we concluded that we were correct in predicting that this change was enhancive. The prediction of the secondary category was also correct.

@@

%%Describe the change 
Version 2.0 of the API was designed to be an event-driven system which makes use of callbacks. To achieve this, the old event system had to be re-designed. The new design is used by all features of API version 2.0. An example of when an event is generated is when movement is detected by the camera.

%%Relate it to Categories
Given that this module was a re-implementation of the previous event system, we concluded that it should be classified mainly as an enhancive change with the secondary category being classified as groomative. Groomative was chosen as the secondary category due to the additional motive of bringing the new functionality into line with the new design.

%%What she said 
The interviewee outlined that the previous version of axevent, event, was deprecated in favour of a re-implemention of the new event system. One reason being that the case company were dissatisfied with certain usability aspects of the API. The module was then completely re-implemented and according to the new system design of version 2.0.

%%Category based on her information
From the discussion in the interview, it was clear that both of our categorisations were correct. The change was primarily enhancive and secondarily groomative.

@@

%%Describe the change 
The module axhttp serves to provide a Common Gateway Interface which allows the camera to dynamically load data to web pages. The previous version of axhttp was re-implemented in version 2.0, to correspond with the new design standards. Additional functionality was also added to the module.

%%Relate it to Categories
This change re-implemented an existing module while adhering to the new design of API version 2.0. The main categorisation was chosen to be enhancive with groomative as the secondary change.

%%What she said 
The interviewee expressed that the main driving force behind this change was to re-implement and add new functionality to the axhttp module, while at the same time, allowing them to the module according to the new design.

%%Category based on her information
Our categorisation of enhancive and groomative was therefore correct for this change. The primary intent was to add functionality while also allowing the re-implementation accommodate the new API design.

@@

%%Describe the change 
The parameter setting module of the API allows data and application settings to be saved so that they are not lost when the device is not running. This module was re-implemented as ax\_parameter with added functionality and conformance to the new design standards of API version 2.0.

Given that this change re-implemented an existing module while also incorporating the new design standards, it has been categorised as being primarily enhancive with a secondary category of groomative.

The interviewee clarified that this change was mainly concerned with re-implementing existing functionality while simultaneously adopting the new standards of version 2.0.

Given the information from the interview, we concluded that this change was correctly categorised; the primary intent was enhancive and the secondary intent was groomative.

@@

%%Describe the change 
The burst mode feature of the API allowed users to buffer a series of images for more exact analysis. This feature was deprecated during the transition between version 1.4 and 2.0 and was not re-implemented.

%%Relate it to Categories
This change presented itself as a clear example of a reductive category of maintenance, as it was simply removed and was not replaced. We also concluded that this change had no secondary category.

%%What she said
The interviewee described that in order to continue using this feature in version 2.0 and in newer versions of the camera hardware, it would have to be re-implemented. An impact assessment was undertaken to determine if the removal of the function would be costly. Ultimately it was deemed appropriate and the function was removed from the API.

%%Category based on her information
Given this information, we can conclude that we correctly categorised this as reductive and that there are indeed no other secondary categories to describe this change.

@@

%%Describe the change

The case company included a selection of example programs to assist developers with creating applications. This trend continued in 2.0, but there were some significant improvements. The number of example programs increased, as did the amount of comments accompanying the code which has lead to an more robust set of examples.

%%Relate it to Categories
We therefore concluded that this change was updative, as the example code was updated to the new API and the new techniques that it utilises. We further decided that this change also had a secondary category of reformative.

%%What she said
From the interview it became clear that the improvement of the example code was an important change between the two versions. The interviewee described how the examples were used frequently by their customers, however, the old examples varied significantly, even between similar tasks. This ultimately meant that the old examples were not satisfactory and new ones needed to be created, preferably conforming to a standard. The new example code explicitly covered the new functionality, focusing on new features and showing how they can be combined to create applications.

%%Category based on her information
With this information we concluded that our categorisation was incorrect. It became clear that this change was in actual fact primarily reformative, given that the changes to the documentation were being performed to meet the needs of the API users. It was also clear that the subcategory should therefore have been updative.

@@

%%Describe the change 
As the API code was updated in version 2.0, the API documentation was updated to correspond with the new code. While there were no substantial changes, we found that the documentation is no longer offered as a PDF but is instead only offered as HTML.

%%Relate it to Categories
We categorised this change as being updative with a reformative secondary category for the reasons that the change only updated the documentation to reflect the recent changes in version 2.0.

%%What she said
This sentiment was supported by the interview. The interviewee agreed with our assessment that the documentation was only changed to be up to date with the current workings of the API and how the new systems interact with each other.

%%Category based on her information
As there was no information to contradict our categorisations we have concluded we were correct in our predictions.

%Re-write this, don't say no information etc

@@

%%Describe the change 
Version 1.4 of the API allowed direct access to the headers of each module available in the API. Version 2.0 changed this by including interfaces through which functionality is accessed. This design change has had an effect on the whole API, as it has changed the structure of modules, but not the functionality of them.

%%Relate it to Categories
We therefore concluded that this change would be best categorised as  groomative. The secondary category was categorised as preventive, as this change could be seen as an attempt to reduce future maintenance.

%%What she said
The interview provided further insight into the reasoning behind this design change. Improvements to the coding standards, the desire for an event driven system and difficulty in using the existing APIs were all contributors to the decision to change the interfaces of the API. Furthermore, the re-design would hopefully avoid future maintenance while allowing the system to be easily extended.

%%Category based on her information
Taking this information into account, we concluded that our categorisation was incorrect. The interview showed that the primary categorisation should have been preventive, given the strong emphasis on avoiding future maintenance and improving ease of use. We further concluded that the secondary change should have been groomative to reflect the intent to make the API more maintainable.

@@

%%Describe the change 
A substantial change included in version 2.0 was the addition of header files, containing enums of error codes. While we were not aware of what error handling was utilised prior to this version, we assumed it had changed, given the inclusion of these new files.

%%Relate it to Categories
Based on this, we categorised this change as being primarily preventive as error handling is preventive by nature. Groomative was decided as the secondary category of maintenance, as improvement of error handling could be regarded as an attempt to improve maintainability.

%%What she said
%% Glib - https://developer.gnome.org/glib/
During the interview, it was made clear that the previous error handling was lacking in expressiveness and usability. Glib was added to version 2.0 and the error handling functionality that Glib provides, was incorporated into the new API design.

%%Category based on her information
We therefore concluded that our categorisation for this change was correct. Indeed, the primary reason for the change was to improve the existing error handling while conforming to the new API standard.

@@

\subsubsection{Continuing Change} %Reviewed by W
The case company described that including new functional content was one of the top priorities for the release of vesrion 2.0. This is supported by the characteristic of the changes identified in this study. 4 of the 13 changes were additions of new modules, and additional functionality was also added to re-implemented modules. It was also mentioned that the API users had previously requested more frequent updates to the API, despite the drawbacks this could cause. One of the main reasons for continuously adapting the API is also to control the way API users use the API. Prior to version 2.0 being released, a number of API users used workarounds to implement applications which contained functionality which was not yet offered from the API. These applications were fully functional, but there existed no standard method of implementation. We therefore conclude that the law of continuing change holds for APIs.

\subsubsection{Increasing Complexity} \label{sec:law2} %Reviewed by W
A majority of the changes introduced in version 2.0 aimed to improve the structure and the way that the API was used. Many of these changes were related to the same design choices, which enforced a certain implementation style on the whole API. One example of this is the change to the interfaces which set a standard for the whole API. During the interview it was also concluded that consistency and following certain standards is an important factor to consider when designing APIs. It was also made clear that a contributing factor for changing certain parts of the API was to bring it into line with the new interface design. The interviewee mentioned that one way of achieving this was to follow guidelines or design best practices, but that the company currently did not make use of these. Even though the importance of maintaining the API in order to reduce complexity was stated, arguments which support that it is of lesser importance than releasing functional content was also presented. One example of where this is evident in the evolution of the API, is in the module that handles the video capturing and video stream. It was decided that this module would not be re-designed to conform with the new interface standard, as the changes to functionality in this module were very minor. Thus, the law of increasing complexity holds for APIs, but there is evidence that suggests that decreasing the complexity is not of the highest priority.

\subsubsection{Self Regulation} \label{sec:law3} %Reviewed by W
The release of version 1.4 did not include any changes related to the API interface. Version 1.0 through to 1.4 strictly added support for additional build target platforms. When comparing the changes introduced in version 2.0 to the ones in previous versions, it is clear that the growth of the API has gradually increased. During the interview it was also mentioned that the future plans for the API did not include any major changes and that only a few minor changes were planned in the near future. The future updates to the API will also be made incrementally and there will not be a big bang update similar to that of version 2.0. This supports that the law of self regulation applies to APIs, assuming that version 2.0 can be considered to be the peak of the normal distribution curve of the APIs growth.

\subsubsection{Conservation of Organisational Stability} %Reviewed by W
The changes to the API have been made very sporadically, where only two new versions have been released since the deployment of version 1.0. As mentioned in Section \ref{sec:law3}, the changes introduced in version 1.4 did not involve any changes to the existing API functions. Consequently, the only significant update to the API was made in the transition to version 2.0. The internal development of the API has also been made sporadically and there have been periods where the API has not been actively developed. This suggests that the law of conservation of organisational stability does not apply to APIs.

\subsubsection{Conservation of Familiarity} %Reviewed by W
The documentation that is included with the API was rigorously updated with the introduction of version 2.0. All functionality which was added or modified in version 2.0 was updated accordingly in the documentation. The same pattern can be seen for the example code, which went through major changes as a result of the changes to the API code. The case company's intent to control the way that their API is used also suggests that the knowledge of how to use the API is of importance, since if the API is used in too many different ways, it will be more difficult for the case company to have relevant documentation and example code. It was also expressed during the interview that further transparency in how the API should be used could be achieved. One example mentioned of how this could be achieved is by improving the expressiveness regarding which API functions are supported on which hardware types. These facts support that the law of conservation of familiarity is true for APIs.

\subsubsection{Continuing Growth} %Reviewed by W
Large parts of the changes between the versions were strictly additions of new functional content. 4 new modules were added and significant additions of new functionality were added to the modules which were re-designed from version 1.4. The interviewee also made it clear that increasing the functionality was one of the main goals with the 2.0 update. This also ties in with the ambition of controlling the way that the API is used, as mentioned previously. It was also mentioned that the case company had received feedback from the API users that more frequent updates to the functional content was desirable, even though this may require them to update their application code. We therefore conclude that the law of continuing growth applies to APIs.

\subsubsection{Declining Quality} %Reviewed by W
The interviews established that a groomative motive was not the primary reason for implementing any change. The changes to the example code and the documentation, as well as the re-designed modules are examples of these, where the main motive behind the change was to either prevent unwanted usage or to increase the functional content. These changes did, however, also increase the quality and were a substantial part of the update. A counterargument to this reasoning can be seen in the module which handles the video data and capturing, which was not changed between the versions, as mentioned in Section \ref{sec:law2}. We therefore settle that the law of declining quality can be applied to APIs, but that there is data that implies that reducing quality is not of high importance.

\subsubsection{Feedback System} %Reviewed by W
One of the future goals of improvement of the case company is to better include the API users and end-users of the applications in their feedback loop. Currently, only the API users are included in this feedback loop and the API architect expressed the desire to also include the end-users. The reasoning for this was to increase the quality, and to be able to deliver content which is useful for the users and which leads to good applications being developed. Prior to version 2.0 being released, an example of how the API users were included in the feedback loop was when they implemented functionality related to sound, despite the fact there was no interface for this.  The experiences of this API user, were later considered when developing the axsound module. One benefit of having a more structured feedback loop, would be that the API developers would have greater awareness of potential errors in the implementation of the API. Another ambition which was mentioned, was to bring scientific theories and best practices from academia into consideration when developing the API. This would aim to complement the current strategy which is largely based around the expertise and experiences of the API architects. Based on these factors and clearly expressed ambitions, we conclude that the law of feedback system is true for APIs.

\section{Discussion}
Our results have been provided valuable insight to the motives behind the evolution of an API. The results show that APIs definitely can be studied from the point of view of old evolution theories. Based on the result, we have formulated the subsequent outcomes.

@@

\noindent
\textbf{Adding functionality is a prime driver of API evolution  }
Overwhelmingly, the primary driving factor behind API evolution was the desire for new API functionality. We found strong evidence to suggest that most, if not all of the additions, re-implementations and functional changes were undertaken to include additional functionality requested both internally and by the API users. This focus upon functionality was also emphasised in part due to a business goal. The more features the API can provide, the more applications developers can create for the cameras. This has a knock-on effect of increasing the appeal and therefore value of the products.

%API users should be included in the design process?
\smallskip \noindent
\textbf{Maintainability does not drive change  } 
While maintainability is important, it is not a primary driving factor for change. We found that where maintainability changed, functionality was also altered. For example, we predicted that the main intents to be enhancive which was correct. Furthermore, of the changes identified, none were considered to be primarily groomative, groomative changes were only identified as sub-category changes. Changes to the example code is further evidence of this. The example code was updated purely to reflect the changes in functionality and not as a groomative effort.

\smallskip \noindent
\textbf{Aiding development of applications  } 
Aiding users of the API to create correct and robust applications is a very important evolution factor for Axis. Changes between the versions brought about improvements to error handling, the example code and the way in which the API utilises interfaces. This, coupled with removal of unwanted functionality and unsafe operations, has allowed the API to grow and become much more user friendly. As mentioned previously, this could increase the amount of applications being created for the camera and thus could increase its value.

\smallskip \noindent
\textbf{Removing functionality  }
Previous literature has stated that deprecations should be favoured over removals from API for many reasons except in specific circumstances. The successful removal of the burst function seems to indicate that this advice is well worth following. Indeed, axis only removed this function from the API after an internal review found that the cost of removing it would outweigh potential drawbacks to keeping it in the API. It is therefore recommended to only remove something if the effects of removing it have been well investigated and that the removal is rather small in scale.

\smallskip \noindent
\textbf{Updative changes do not drive change  }
Changes to documentation and example code are important, especially if the API is drastically changing between versions. However, these updative changes are drivers themselves, rather, they make up part of a larger change as a whole. 
%^Probably needs more to it?

\smallskip \noindent
\textbf{Hardware was not a driving factor  }
Interestingly, we found that hardware was not a direct driver of change. Initially we had thought that it would be one of the strongest drivers of change, for example, new versions of hardware become available or new hardware is added to the camera requiring something in the API to access it. This does seem to suggest that the hardware for the camera is sufficiently stable such that it does not require changes and that those changes do not need to be reflected in new API versions.

>>>>

%%Add percentages for both main and sub and specify which is for which
After cross-examining our results with the interviewee and their responses, we found that roughly 78\% of our categorisation matched the actual categorisation of the changes.

@@

\subsubsection{axptz}
%Add info about this is one of the added changes XXX (do that for all)
The Pan, Tilt and Zoom (ptz) module concerns functionality related to manoeuvring the camera, both mechanically and digitally. 
%%Relate it to Categories
We concluded that this change was enhancive as it is a brand-new feature of the API. There was no secondary motive behind this change, as the change was purely enhancive. 
%%What she said
The interviewee described how this change and the other changes classified as additions were all included in the new API at the request of API users. Prior to the change being introduced, there was no API functionality for controlling the camera and as such, it was a high priority addition to the new API. 
%%Category based on her information
Based on the interviewees' comments, we concluded that we had correctly identified the change as being enhancive. Furthermore, we were also correct in our prediction that there was no secondary category.

\subsubsection{axaudio}
%%Describe the change 
The most important functionality of the camera is to capture and analyse video data, however, some the API users require the possibility to perform analysis upon audio data. The addition of axaudio allows API users to perform audio analysis upon audio streams captured by cameras. 
%%Relate it to Categories
This change was determined to be enhancive, as it was an addition of a new feature to the API. We found no secondary motive of the change. 
%%What she said
Audio analysis had not previously been possible on Axis devices. During the interview, it was revealed that an API user had managed to implement their own audio analysis using version 1.4 of the API. Given the demand for audio analysis, it was included in version 2.0 of the API.
%%Category based on her information
Based on the the interviewee's reasoning, we concluded that we had successfully determined that the change was enhancive. The addition of axaudio had no secondary motive.

@@

%%Describe the change 
Prior to the newest version of the API being released, the sole alternative for storage was to use the Video Management System. Version 2.0 included the possibility of alternative storage solutions, by allowing to the camera to interface with external storage devices, such as SD cards. 
%Relate it to categories
Enhancive was chosen to be the most applicable category of change, given that axstorage was a new addition in version 2.0. No secondary motive was identified. 
%%What she said
The limited storage options for the cameras was the reason that increased functionality was strongly requested by API users and as such axstorage was included in version 2.0 of the API.
%%Category based on her information
The information gathered from the interview supported that we were correct in our categorisation of this change. No secondary category was established.

@@

%%Describe the change 
The addition of axserialport allows API users to connect various devices to their cameras by through the serial port. This allows users to integrate multiple devices to create their own unique solutions. In version 1.4 of the API, there was no possibility to connect external devices to the camera.
%%Relate it to Categories
The addition of this module was determined to be an enhancive change, as it involved the addition of a new feature. No relevant secondary category were identified.
%%What she said
The addition of axserialport was requested by API users. A possible use case, which was given during the interview involved a ticketing system which allows for joint analysis of data from both devices. As such it was included in version 2.0, to meet the demand of the API users.
%%Category based on her information
Given this information, we concluded that we were correct in predicting that this change was enhancive. The prediction of the secondary category was also correct.

@@

%%Describe the change 
Version 2.0 of the API was designed to be an event-driven system which makes use of callbacks. To achieve this, the old event system had to be re-designed. The new design is used by all features of API version 2.0. An example of when an event is generated is when movement is detected by the camera. 
%%Relate it to Categories
Given that this module was a re-implementation of the previous event system, we concluded that it should be classified mainly as an enhancive change with the secondary category being classified as groomative. Groomative was chosen as the secondary category due to the additional motive of bringing the new functionality into line with the new design.
%%What she said 
The interviewee outlined that the previous version of axevent, event, was deprecated in favour of a re-implemention of the new event system. One reason being that the case company were dissatisfied with certain usability aspects of the API. The module was then completely re-implemented and according to the new system design of version 2.0.
%%Category based on her information
From the discussion in the interview, it was clear that both of our categorisations were correct. The change was primarily enhancive and secondarily groomative.

@@

%%Describe the change 
The module axhttp serves to provide a Common Gateway Interface which allows the camera to dynamically load data to web pages. The previous version of axhttp was re-implemented in version 2.0, to correspond with the new design standards. Additional functionality was also added to the module.
%%Relate it to Categories
This change re-implemented an existing module while adhering to the new design of API version 2.0. The main categorisation was chosen to be enhancive with groomative as the secondary change.
%%What she said 
The interviewee expressed that the main driving force behind this change was to re-implement and add new functionality to the axhttp module, while at the same time, allowing them to the module according to the new design.
%%Category based on her information
Our categorisation of enhancive and groomative was therefore correct for this change. The primary intent was to add functionality while also allowing the re-implementation accommodate the new API design.

@@

%%Describe the change 
The parameter setting module of the API allows data and application settings to be saved so that they are not lost when the device is not running. This module was re-implemented as ax\_parameter with added functionality and conformance to the new design standards of API version 2.0.
%2
Given that this change re-implemented an existing module while also incorporating the new design standards, it has been categorised as being primarily enhancive with a secondary category of groomative.
%3
The interviewee clarified that this change was mainly concerned with re-implementing existing functionality while simultaneously adopting the new standards of version 2.0. 
%4
Given the information from the interview, we concluded that this change was correctly categorised; the primary intent was enhancive and the secondary intent was groomative.

@@

%%Describe the change 
The burst mode feature of the API allowed users to buffer a series of images for more exact analysis. This feature was deprecated during the transition between version 1.4 and 2.0 and was not re-implemented.
%%Relate it to Categories
This change presented itself as a clear example of a reductive category of maintenance, as it was simply removed and was not replaced. We also concluded that this change had no secondary category.
%%What she said
The interviewee described that in order to continue using this feature in version 2.0 and in newer versions of the camera hardware, it would have to be re-implemented. An impact assessment was undertaken to determine if the removal of the function would be costly. Ultimately it was deemed appropriate and the function was removed from the API.
%%Category based on her information
Given this information, we can conclude that we correctly categorised this as reductive and that there are indeed no other secondary categories to describe this change.

@@

%%Describe the change 
The case company included a selection of example programs to assist developers with creating applications. This trend continued in 2.0, but there were some significant improvements. The number of example programs increased, as did the amount of comments accompanying the code which has lead to an more robust set of examples.
%%Relate it to Categories
We therefore concluded that this change was updative, as the example code was updated to the new API and the new techniques that it utilises. We further decided that this change also had a secondary category of reformative.
%%What she said
From the interview it became clear that the improvement of the example code was an important change between the two versions. The interviewee described how the examples were used frequently by their customers, however, the old examples varied significantly, even between similar tasks. This ultimately meant that the old examples were not satisfactory and new ones needed to be created, preferably conforming to a standard. The new example code explicitly covered the new functionality, focusing on new features and showing how they can be combined to create applications.
%%Category based on her information
With this information we concluded that our categorisation was incorrect. It became clear that this change was in actual fact primarily reformative, given that the changes to the documentation were being performed to meet the needs of the API users. It was also clear that the subcategory should therefore have been updative.

@@

%%Describe the change 
As the API code was updated in version 2.0, the API documentation was updated to correspond with the new code. While there were no substantial changes, we found that the documentation is no longer offered as a PDF but is instead only offered as HTML. 
%%Relate it to Categories
We categorised this change as being updative with a reformative secondary category for the reasons that the change only updated the documentation to reflect the recent changes in version 2.0.
%%What she said
This sentiment was supported by the interview. The interviewee agreed with our assessment that the documentation was only changed to be up to date with the current workings of the API and how the new systems interact with each other.
%%Category based on her information
As there was no information to contradict our categorisations we have concluded we were correct in our predictions.

%!!!Re-write this, don't say no information etc

@@

%%Describe the change 
Version 1.4 of the API allowed direct access to the headers of each module available in the API. Version 2.0 changed this by including interfaces through which functionality is accessed. This design change has had an effect on the whole API, as it has changed the structure of modules, but not the functionality of them.
%%Relate it to Categories
We therefore concluded that this change would be best categorised as  groomative. The secondary category was categorised as preventive, as this change could be seen as an attempt to reduce future maintenance. 
%%What she said
The interview provided further insight into the reasoning behind this design change. Improvements to the coding standards, the desire for an event driven system and difficulty in using the existing APIs were all contributors to the decision to change the interfaces of the API. Furthermore, the re-design would hopefully avoid future maintenance while allowing the system to be easily extended. 
%%Category based on her information
Taking this information into account, we concluded that our categorisation was incorrect. The interview showed that the primary categorisation should have been preventive, given the strong emphasis on avoiding future maintenance and improving ease of use. We further concluded that the secondary change should have been groomative to reflect the intent to make the API more maintainable.

@@

%%Describe the change 
A substantial change included in version 2.0 was the addition of header files, containing enums of error codes. While we were not aware of what error handling was utilised prior to this version, we assumed it had changed, given the inclusion of these new files.
%%Relate it to Categories
Based on this, we categorised this change as being primarily preventive as error handling is preventive by nature. Groomative was decided as the secondary category of maintenance, as improvement of error handling could be regarded as an attempt to improve maintainability. 
%%What she said
%% Glib - https://developer.gnome.org/glib/
During the interview, it was made clear that the previous error handling was lacking in expressiveness and usability. Glib was added to version 2.0 and the error handling functionality that Glib provides, was incorporated into the new API design.
%%Category based on her information
We therefore concluded that our categorisation for this change was correct. Indeed, the primary reason for the change was to improve the existing error handling while conforming to the new API standard.

@@

\subsubsection{Continuing Change} %Reviewed by WJ
The case company described that including new functional content was one of the top priorities for the release of vesrion 2.0. This is supported by the characteristic of the changes identified in this study. 4 of the 13 changes were additions of new modules, and additional functionality was also added to re-implemented modules. It was also mentioned that the API users had previously requested more frequent updates to the API, despite the drawbacks this could cause. One of the main reasons for continuously adapting the API is also to control the way API users use the API. Prior to version 2.0 being released, a number of API users used workarounds to implement applications which contained functionality which was not yet offered from the API. These applications were fully functional, but there existed no standard method of implementation. We therefore conclude that the law of continuing change holds for APIs.

\subsubsection{Increasing Complexity} \label{sec:law2} %Reviewed by WJ
A majority of the changes introduced in version 2.0 aimed to improve the structure and the way that the API was used. Many of these changes were related to the same design choices, which enforced a certain implementation style on the whole API. One example of this is the change to the interfaces which set a standard for the whole API. During the interview it was also concluded that consistency and following certain standards is an important factor to consider when designing APIs. It was also made clear that a contributing factor for changing certain parts of the API was to bring it into line with the new interface design. The interviewee mentioned that one way of achieving this was to follow guidelines or design best practices, but that the company currently did not make use of these. Even though the importance of maintaining the API in order to reduce complexity was stated, arguments which support that it is of lesser importance than releasing functional content was also presented. One example of where this is evident in the evolution of the API, is in the module that handles the video capturing and video stream. It was decided that this module would not be re-designed to conform with the new interface standard, as the changes to functionality in this module were very minor. Thus, the law of increasing complexity holds for APIs, but there is evidence that suggests that decreasing the complexity is not of the highest priority.

\subsubsection{Self Regulation} \label{sec:law3} %Reviewed by WJ
The release of version 1.4 did not include any changes related to the API interface. Version 1.0 through to 1.4 strictly added support for additional build target platforms. When comparing the changes introduced in version 2.0 to the ones in previous versions, it is clear that the growth of the API has gradually increased. During the interview it was also mentioned that the future plans for the API did not include any major changes and that only a few minor changes were planned in the near future. The future updates to the API will also be made incrementally and there will not be a big bang update similar to that of version 2.0. This supports that the law of self regulation applies to APIs, assuming that version 2.0 can be considered to be the peak of the normal distribution curve of the APIs growth.

\subsubsection{Conservation of Organisational Stability} %Reviewed by WJ
The changes to the API have been made very sporadically, where only two new versions have been released since the deployment of version 1.0. As mentioned in Section \ref{sec:law3}, the changes introduced in version 1.4 did not involve any changes to the existing API functions. Consequently, the only significant update to the API was made in the transition to version 2.0. The internal development of the API has also been made sporadically and there have been periods where the API has not been actively developed. This suggests that the law of conservation of organisational stability does not apply to APIs.

\subsubsection{Conservation of Familiarity} %Reviewed by WJ
The documentation that is included with the API was rigorously updated with the introduction of version 2.0. All functionality which was added or modified in version 2.0 was updated accordingly in the documentation. The same pattern can be seen for the example code, which went through major changes as a result of the changes to the API code. The case company's intent to control the way that their API is used also suggests that the knowledge of how to use the API is of importance, since if the API is used in too many different ways, it will be more difficult for the case company to have relevant documentation and example code. It was also expressed during the interview that further transparency in how the API should be used could be achieved. One example mentioned of how this could be achieved is by improving the expressiveness regarding which API functions are supported on which hardware types. These facts support that the law of conservation of familiarity is true for APIs.

\subsubsection{Continuing Growth} %Reviewed by WJ
Large parts of the changes between the versions were strictly additions of new functional content. 4 new modules were added and significant additions of new functionality were added to the modules which were re-designed from version 1.4. The interviewee also made it clear that increasing the functionality was one of the main goals with the 2.0 update. This also ties in with the ambition of controlling the way that the API is used, as mentioned previously. It was also mentioned that the case company had received feedback from the API users that more frequent updates to the functional content was desirable, even though this may require them to update their application code. We therefore conclude that the law of continuing growth applies to APIs.

\subsubsection{Declining Quality} %Reviewed by WJ
The interviews established that a groomative motive was not the primary reason for implementing any change. The changes to the example code and the documentation, as well as the re-designed modules are examples of these, where the main motive behind the change was to either prevent unwanted usage or to increase the functional content. These changes did, however, also increase the quality and were a substantial part of the update. A counterargument to this reasoning can be seen in the module which handles the video data and capturing, which was not changed between the versions, as mentioned in Section \ref{sec:law2}. We therefore settle that the law of declining quality can be applied to APIs, but that there is data that implies that reducing quality is not of high importance.

\subsubsection{Feedback System} %Reviewed by WJ
One of the future goals of improvement of the case company is to better include the API users and end-users of the applications in their feedback loop. Currently, only the API users are included in this feedback loop and the API architect expressed the desire to also include the end-users. The reasoning for this was to increase the quality, and to be able to deliver content which is useful for the users and which leads to good applications being developed. Prior to version 2.0 being released, an example of how the API users were included in the feedback loop was when they implemented functionality related to sound, despite the fact there was no interface for this.  The experiences of this API user, were later considered when developing the axsound module. One benefit of having a more structured feedback loop, would be that the API developers would have greater awareness of potential errors in the implementation of the API. Another ambition which was mentioned, was to bring scientific theories and best practices from academia into consideration when developing the API. This would aim to complement the current strategy which is largely based around the expertise and experiences of the API architects. Based on these factors and clearly expressed ambitions, we conclude that the law of feedback system is true for APIs.

\section{Discussion} %Reviewed by W
Our results have been provided valuable insight to the motives behind the evolution of an API. The results show that APIs definitely can be studied from the point of view of long-established software evolution theories. Based on the result, we have formulated the subsequent outcomes.

@@

\noindent
\textbf{Adding functionality is a prime driver of API evolution  } %Reviewed by W
Overwhelmingly, the primary driving factor behind API evolution was the desire for new API functionality. We found that a vast majority of the changes either added functionality or were direct effects of functionality being added. The main driver for these changes was the pressure from the API users, but pressure from the case company developers which utilised the API also contributed. This focus upon functionality was also emphasised in part due to a business goal ( "in part due to, sounds weird to me") . The more features the API can provide, the more applications developers can create for the cameras. This has a domino effect of increasing the appeal and therefore value of the products.

%API users should be included in the design process? 
\smallskip \noindent 
\textbf{Maintainability does not drive change  } %Reviewed by W 
While maintainability is important, it is not a primary driving factor for change. We found that where maintainability changed, functionality was also altered. For example, we predicted that the main intents to be enhancive which was correct. ( " Need to explain what was enhancive") Furthermore, of the changes identified, none were considered to be primarily groomative. Groomative motives were only found to be secondary, and never drove the change. Changes to the example code is further evidence of this. The example code was updated purely to reflect the changes in functionality and not as a groomative effort.

\smallskip \noindent
\textbf{Aiding development of applications  } %Reviewed by W
Aiding ( "minor detail: is supporting better here?") users of the API to create correct and robust applications is a very important evolution factor for the case company. Changes between the versions brought about ( 're-word'?) improvements to error handling, the example code and the way in which the API utilises interfaces. This, coupled with removal of unwanted functionality and unsafe operations, has allowed the API to grow and become much more user-friendly. As mentioned previously, this could increase the amount of applications being created for the camera and in turn could increase its value ( "change to' : increase the value of them ?").

\smallskip \noindent
\textbf{Removing functionality  } %Reviewed by W
Previous literature has stated that deprecations should be favoured over removals from API for many reasons, except in specific circumstances. The successful removal of the burst function seems to indicate that this advice is well worth following ( "well worth following reads weirdly") ( 'this sentence also contradicts the previous one'). Indeed, Axis only removed this function from the API after an internal review found that the cost of removing it would outweigh potential drawbacks to ( 'of???') keeping it in the API. ( 'this sentence is weird', "and remove the indeed") It is ( 'change to We therefore recommend') therefore recommended to only remove something if the effects of removing it have been well investigated and that the removal is rather small in scale. ( "instead of saying  small in scale i think we should explain that they should be made in increments and only a few at once")

\smallskip \noindent
\textbf{Updative changes do not drive change  } %Reviewed by W
Changes to documentation and example code are important, especially if the API is drastically changing between versions. However, these updative changes are drivers themselves, rather, they make up part of a larger change as a whole. ( "this whole section has to be re-written. We can't refer to them as updative changes since the nature of updative changes are reactive. instead refer to them as changes to !code e.g. documentation, example code etc.")
%^Probably needs more to it?
